\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-06 europa-basics.tex
updates
CSP  background section looking close to complete now.

\eu is a versatile planning framework which can be used in a range of
problem-solving methods.

\paragraph \texttt{Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N$x$N$ chessboard so no queens attack the other. Fig.
\ref{fig:nqueens-1} shows an example of a random positioning of queens
on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
are highlighted in red.

@@

\paragraph \texttt{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \comment{citations needed}, a project
consisting of a set of activities must be scheduled in a way that
satisfies minimum and/or maximum temporal separation constraints. The
activity schedule must also respect fixed limits on the availability of
resources required to perform each activity. In addition to satisfying
temporal and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

@@

\paragraph \texttt{Planning}: In the Shopping Agent Problem
\cite{russelnorvig} an agent needs to purchase a set of products (milk,
drill, etc) that are available at specific locations (supermarket,
hardware store, etc), the agent is subject to temporal (must complete
tasks by specific deadlines) and resource (fuel, carrying capacity, etc)
constraints. The agent needs to figure out what actions need to be
performed to find and acquire the required items, as well as when to
perform each of those actions. Fig. \ref{fig:shopping-1} shows a
solution produced by \eu for such a problem instance.

@@

As the examples above show, a planning problem (actions to achieve a
goal) may embed a scheduling problem (what resources are necessary to
achieve stated goals) and both planning and scheduling could embed a
Constraint Satisfaction problem. % since temporal, resource and  other
% kinds of constraints must be respected in most real life problems.
The relationships between Planning, Scheduling and Constraint
Satisfaction have been examined \cite{smith00} and have lead to use of
constraint reasoning in \eu as its innermost building block. % Below we
% examine the main ideas from Constraint Satisfaction Programming, and
% from Constraint Based Planning and Scheduling that constitute
% \texttt{EUROPA}$_2$'s theoretical underpinnings.

@@

To solve a problem, CP techniques use logical inference to perform two
important operations:

@@

CP is normally implemented as part of a programming language and
constraints are normally represented as objects [PUG95]
\comment{citation}. Any constraint can be introduced and, as long as the
bounds propagation and domain reduction protocols specified by the host
CP system are enforced, it will be indistinguishable from any other
``primitive'' CP constraint, such as $\leq$ or $\geq$. In theory,
solving a CSP problem is NP-Hard \cite{ghallab04}, but often very
efficient in practice using a number of algorithms and techniques that
are well understood.

\comment{TODO: talk about arc consistency algorithms, computational
  complexity and solvers.}

@@

\subsection{EUROPA's Plan Representation}
\label{sec:europa:pr}

EUROPA is an implementation of the CAIP framework explained in the previous section

Variables

Values that need to be represented to describe the problem domain and over which we may want to specify constraints.

Objects

The things we wish to describe and refer to in a domain are considered Objects. As in the case with object-oriented analysis and design, one can seek out the nouns in any domain description to find likely objects. In a Mars rover example, we might consider the satellite, the camera, and the attitude controller (which orients the robot for taking pictures) to be objects. Objects have state and behavior. For example, a camera can be:

@@

An object is an instance of a class. In EUROPA we model using the abstraction of a class to speak about all instances having certain properties of state and behavior. In order to describe such state and behavior we build on the formalism of first order logic.

Temporally Qualified Predicates and Actions

A predicate defines a relation between objects and properties. In EUROPA, we define such relations between variables whose domains are sets of objects and sets of properties. For example, we might use a predicate Pointing(a,p) to indicate that a rover's attitude controller a is pointing at position p. Note that a is a variable which may have a number of possible values in a problem with multiple satellites. Similarly, p is a variable whose values are the set of possible positions. In a grounded plan, single values will be specified for each variable as we saw in the case of a solved Constraint Satisfaction Problem.

In general, is is not sufficient to state that a predicate is true without giving it some temporal extent over which it holds. Predicates that are always true can be thought to hold from the beginning to the end of time. However, in practice, the temporal extent of interest must be defined with timepoints to represent its start and end. So we might want to write Pointing(a,s,e,p) to indicate that the attitude controller a is pointing at position p from time s to time e. In fact, this pattern of using such predicates to describe both state and behavior of objects is so prevalent in EUROPA that we have introduced a special construct called a Token which has the built in variables to indicate the object to which the statement principally applies and the timepoints over which it holds. In EUROPA, all predicate instances are Tokens.

Tokens

@@

    start: The beginning of the temporal extent over which the predicate is defined.
    end: The end of the temporal extent over which the predicate is defined.
    duration: The constraints start + duration = end is enforced automatically.
    object: made. The set of objects to which a token might apply. In a grounded plan each Token applies to a specific Object, reflecting the intuition that we are using Tokens to describe some aspect of an Object (i.e. its state or behavior) in time. However, in a partial plan, the commitment to a specific object may not yet have been made.
    state: Tokens can be ACTIVE, INACTIVE, MERGED, or REJECTED. The state variable captures the token's current state and its reachable states through further restriction. See the Token State Model discussion below for details.

Objects (continued)

@@

    Static Facts: Classes without predicates lead to Objects without Tokens. This arises where an objects state or behavior is independent of time. For example, a domain may have a set of locations and/or paths for which there is nothing more to say than that they exist.
    Timelines: Often objects in a domain must be described by exactly one token for every given timepoint in the plan. Such objects are so common that we provide a special construct in EUROPA to extend these semantics to derived classes. Any instances of a class derived from a Timeline will induce ordering requirements among its tokens in order to ensure no temporal overlap may occur among them. See below for further discussion.
    Resources: Metric resources, e.g. the energy of a battery or the capacity of a cargo hold, are objects with an explicit quantitative state in time and with a circumscribed range of changes that can occur to impact that state i.e. produce, consume, use, change. These changes are captured as tokens. Resources are such a common requirement for EUROPA users that special constructs are also provided for them. Instances of classes derived from a Resource will induce ordering requirements on their Tokens in order to ensure that the level of the resource remains within specified limits.

@@

The state of a token is embodied with a 5th and final built-in variable referred to as the state variable. The planner states are values in the domain of this variable {MERGED, ACTIVE, REJECTED}. The INACTIVE state is captured by the variable being unbound. If a value is removed from the domain, that state will not be reachable. For example, when a slave is created via a rule, the REJECTED value is removed from its state variable.
Summary

This page has presented the main elements of the EUROPA planning approach. They are:

    Temporally scoped predicates and actions (a.k.a. Tokens) to represent states and actions in time. Note that tokens do not discriminate between state and action.
    Constraints to describe relationships among tokens. This provides an expressive method of describing interactions among tokens in the context of Temporal Planning.
    Timelines as a concise abstraction to express the evolution of state and behavior for a system component. It provides semantics of mutual exclusion sequencing in time. Other core abstractions are available in EUROPA for handling metric resources.
    Domain rules to describe internal and external relationships on and between tokens respectively. Rules are applied to active tokens (master tokens) referred to as masters and typically produce inactive tokens (slaves) which are a key vehicle through which the planning process evolves. The rule structure in EUROPA differs from the more restrictive commitment to preconditions and effects in classical planning which prohibits durative actions and disembodied effects (i.e. effects which may not occur until some temporal distance after the end of the action).
    Token States which are the basis of planning operations on a partial plan.

>>>>

\eu is a versatile planning framework, let's start by looking at some examples of the main problem types that it can tackle.

\paragraph \textit{Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N$x$N$ chessboard so no queens attack the other. Fig.
\ref{fig:nqueens-1} shows an example of a random positioning of queens
on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
are highlighted in red.

@@

\paragraph \textit{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \comment{citations needed}, a project
consisting of a set of activities must be scheduled in a way that
satisfies minimum and/or maximum temporal separation constraints. The
activity schedule must also respect fixed limits on the availability of
resources required to perform each activity. In addition to satisfying
temporal and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

@@

\paragraph \textit{Planning}: In the Shopping Agent Problem
\cite{russelnorvig} an agent needs to purchase a set of products (milk,
drill, etc) that are available at specific locations (supermarket,
hardware store, etc), the agent is subject to temporal (must complete
tasks by specific deadlines) and resource (fuel, carrying capacity, etc)
constraints. The agent needs to figure out what actions need to be
performed to find and acquire the required items, as well as when to
perform each of those actions. Fig. \ref{fig:shopping-1} shows a
solution produced by \eu for such a problem instance.

@@

As the examples above show, a Planning problem (actions to achieve a goal) may embed a Scheduling problem (what resources are necessary to achieve stated goals) and both Planning and Scheduling may embed a Constraint Satisfaction problem. % since temporal, resource and  other
% kinds of constraints must be respected in most real life problems.
The relationships between Planning, Scheduling and Constraint
Satisfaction have been examined \cite{smith00} and have lead to use of
constraint reasoning in \eu as its innermost building block.

@@

To solve a problem, CP techniques use logical inference to perform Constraint Propagation, which consists of two
operations:

@@

A set of constraints and variables that describe a problem domain are typically represented as a Constraint Network, where the variables are nodes , and constraints are arcs between the variables. Having this network representation in mind, Arc Consistency for a a pair of variables x,y is achieved if for every value in x's domain, there is a value in the y's domain that is consistent with all the constraints that connect x and y. 
A naive approach to ensuring arc consistency consists of cycling through all the variable pairs and performing constraint propagation until there are no variable domain changes, the AC-3 algorithm \comment {citation [MAC77]} improves over the naive approach by ignoring constraints whose variables were not affected since the last iteration. AC-3 is very popular in CP implementations.

Arc Consistency and other, more sophisticated consistency algorithms \comment {citation CP Handbook} can be used to efficiently prove that a CP problem is unsatisfiable, however, they generally can't prove that a CP problem is satisfiable and provide specific variable assignments that constitute solutions. For that, search algorithms like global backtracking or local search are used \comment {citation CP Handbook}, which use consistency as an efficient way to prune the search space.

In theory, solving a CP problem is NP-Hard \cite{ghallab04}, but often very efficient in practice using a number of consistency and search algorithms that are well understood.

CP is usually implemented as part of a programming language and constraints are usually represented as objects \cite{puget95}. Any constraint can be introduced and, as long as the constraint propagation protocols specified by the hostCP system are enforced, it will be indistinguishable from any other ``primitive'' CP constraint, such as $\leq$ or $\geq$.

@@

\subsection{\eu Plan Representation}
\label{sec:europa:pr}

\eu is an implementation of the CAIP framework.

\textit{Variables} Values that need to be represented to describe the problem domain and over which we may want to specify constraints.

\textit{Objects} The things we wish to describe and refer to in a domain are considered
Objects. As in the case with object-oriented analysis and design, one
can seek out the nouns in any domain description to find likely
objects. In a Mars rover example, we might consider the satellite, the
camera, and the attitude controller (which orients the robot for
taking pictures) to be objects. Objects have state and behavior. For
example, a camera can be:

@@

    An object is an instance of a class. In \eu we model using the
    abstraction of a class to speak about all instances having certain
    properties of state and behavior. In order to describe such state
    and behavior we build on the formalism of first order logic.

\textit{Tokens}

A predicate defines a relation between objects and properties. In \eu,
we define such relations between variables whose domains are sets of
objects and sets of properties. For example, we might use a predicate
\texttt{Pointing($a,p$)} to indicate that a rover's attitude controller
$a$ is pointing at position $p$. Note that $a$ is a variable which may
have a number of possible values in a problem with multiple
satellites. Similarly, $p$ is a variable whose values are the set of
possible positions. In a grounded plan, single values will be
specified for each variable as we saw in the case of a solved
Constraint Satisfaction Problem.

In general, it is not sufficient to state that a predicate is true
without giving it some temporal extent over which it holds. Predicates
that are always true can be thought to hold from the beginning to the
end of time. However, in practice, the temporal extent of interest
must be defined with timepoints to represent its start and end. So we
might want to write \texttt{Pointing($a,s,e,p$)} to indicate that the
attitude controller a is pointing at position $p$ from time $s$ to
time $e$. In fact, this pattern of using such predicates to describe
both state and behavior of objects is so prevalent in \eu that we have
introduced a special construct called a Token which has the built in
variables to indicate the object to which the statement principally
applies and the timepoints over which it holds. In \eu, all predicate
instances are Tokens.

TODO: talk about actions

@@

start: The beginning of the temporal extent over which the predicate
is defined.  end: The end of the temporal extent over which the
predicate is defined.  duration: The constraints start + duration =
end is enforced automatically.  object: made. The set of objects to
which a token might apply. In a grounded plan each Token applies to a
specific Object, reflecting the intuition that we are using Tokens to
describe some aspect of an Object (i.e. its state or behavior) in
time. However, in a partial plan, the commitment to a specific object
may not yet have been made.

state: Tokens can be ACTIVE, INACTIVE, MERGED, or REJECTED. The state
variable captures the token's current state and its reachable states
through further restriction. See the Token State Model discussion
below for details.

\textit{Objects (continued)}

@@

Static Facts: Classes without predicates lead to Objects without
Tokens. This arises where an objects state or behavior is independent
of time. For example, a domain may have a set of locations and/or
paths for which there is nothing more to say than that they exist.

Timelines: Often objects in a domain must be described by exactly one
token for every given timepoint in the plan. Such objects are so
common that we provide a special construct in \eu to extend these
semantics to derived classes. Any instances of a class derived from a
Timeline will induce ordering requirements among its tokens in order
to ensure no temporal overlap may occur among them. See below for
further discussion.

Resources: Metric resources, e.g. the energy of a battery or the
capacity of a cargo hold, are objects with an explicit quantitative
state in time and with a circumscribed range of changes that can occur
to impact that state i.e. produce, consume, use, change. These changes
are captured as tokens. Resources are such a common requirement for
\eu users that special constructs are also provided for
them. Instances of classes derived from a Resource will induce
ordering requirements on their Tokens in order to ensure that the
level of the resource remains within specified limits.

@@

The state of a token is embodied with a 5th and final built-in
variable referred to as the state variable. The planner states are
values in the domain of this variable {MERGED, ACTIVE, REJECTED}. The
INACTIVE state is captured by the variable being unbound. If a value
is removed from the domain, that state will not be reachable. For
example, when a slave is created via a rule, the REJECTED value is
removed from its state variable.

\textit{Summary}

This has presented the main elements of the EUROPA planning approach. They are:

\item Temporally scoped predicates and actions (a.k.a. Tokens) to represent
states and actions in time. Note that tokens do not discriminate
between state and action.

\item Token States which are the basis of planning operations on a partial
plan.

\item Constraints to describe relationships among tokens. This provides an
expressive method of describing interactions among tokens in the
context of Temporal Planning.

\item Timelines as a concise abstraction to express the evolution of state
and behavior for a system component. It provides semantics of mutual
exclusion sequencing in time. Other core abstractions are available in
\eu for handling metric resources.

\item Domain rules to describe internal and external relationships on and
between tokens respectively. Rules are applied to active tokens
(master tokens) referred to as masters and typically produce inactive
tokens (slaves) which are a key vehicle through which the planning
process evolves. The rule structure in \eu differs from the more
restrictive commitment to preconditions and effects in classical
planning which prohibits durative actions and disembodied effects
(i.e. effects which may not occur until some temporal distance after
the end of the action).
